### 2019-06-14 paper.tex
Comment out future sections

\documentclass[conference]{classes/IEEEtran}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\title{Optimizing TensorFlow's Memory Swapping by Prioritizing Oldest Tensors}

\author{\IEEEauthorblockN{Devin Alvaro}
\IEEEauthorblockA{\textit{School of Electrical Engineering and Informatics} \\
\textit{Bandung Institute of Technology}\\
Bandung, Indonesia \\
devin.alvaro@gmail.com}
\and
\IEEEauthorblockN{Achmad Imam Kistijantoro}
\IEEEauthorblockA{\textit{School of Electrical Engineering and Informatics} \\
\textit{Bandung Institute of Technology}\\
Bandung, Indonesia \\
imam@informatika.org}
}

\maketitle

\begin{abstract}
    The ever-increasing sizes of deep learning models and datasets used increase the need for memory, as insufficient memory may abort a training process. Adding to this problem, deep learning trainings tend to use GPUs over CPUs for better training speed, where in general a GPU has significantly less memory than its CPU counterpart.

    One solution is \textit{memory swapping}, a way to free some memory by temporarily moving data to another memory pool, in this case moving data from a GPU memory to a CPU memory. However, since moving data takes time---the larger the data the more the time---using memory swapping in a training can significantly increase the training duration. Therefore, an optimal memory swapping has to be selective on how much and which data to swap.

    TensorFlow, a machine learning framework, features memory swapping that, based on our analysis over its implementation, can be optimized to reduce the increase to training duration it causes. The optimization is done is by prioritizing oldest tensors (data representation in TensorFlow) which, because of a certain backpropagation property, allows the asynchronicity of program execution to increase, ultimately reducing training duration.

    Based on our experiments, the optimization reduces up to around 3\% of training duration on certain cases.
\end{abstract}

\begin{IEEEkeywords}
memory swapping, TensorFlow, asynchronous, backpropagation
\end{IEEEkeywords}

% \section{Introduction}

% \section{Related Works}

% \section{Memory Swapping in TensorFlow}

% \section{Optimization Design and Implementation}

% \section{Experiments}

% \section{Conclusion}

% \begin{thebibliography}{00}
% \bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
% \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
% \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
% \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
% \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
% \bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
% \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
% \end{thebibliography}

\end{document}